---
title: "Chocolate Bar Rating Analysis"
author: "Joel Nyongesa"
date: "2023-02-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction.
In this project, I will be performing data analysis the `flavors of cacao` dataset in R. The dataset is available on Kaggle, via [this link](https://www.kaggle.com/datasets/rtatman/chocolate-bar-ratings). I will be using the 6 phases of data analysis:


- Ask.
- Prepare.
- Process.
- Analyze.
- Share.
- Act.

## 1. Ask.
Chocolate is one of the most popular candies in the world. Each year, residents of the United States collectively eat more than 2.8 billions pounds. This analysis aims to answer a few business questions to determine the best chocolate producing countries, and the most preferred chocolate types.

The objective of this analysis is to answer the following questions:

- What coffee manufacturing companies received the highest ratings, and where are the companies located?
- Which year had the maximum ratings ratings?
- Which broad bean has the highest rating?

## 2. Prepare.
The data was stored in the form of a csv file. The data had nine columns which are:


- Company.
- Specific ben origin.
- Ref.
- Review.
- Cocoa.
- Company location.
- Rating.
- Bean.
- Broad bean.

The dataset is licenced under [CCO public domain](https://creativecommons.org/publicdomain/zero/1.0/). 

## 3. Process.
Data analysis will be done using R, with documentation done using R Markdown.
The data cleaning process is as follows:

- First loading the dataset.
- Renaming the columns.
- Changing the data types to a suitable format.

```{r loading the required packages}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
```

```{r loading the dataset}
cocoa_flavors <- read.csv("flavors_of_cacao.csv")
```

Now let us inspect the dataset, viewing the structure, first few rows, column names and making any changes that will be necessary.
```{r first few rows}
head(cocoa_flavors)
```

The structure of the dataset.
```{r structure of the dataset}
str(cocoa_flavors)
```

The column names
```{r column names}
colnames(cocoa_flavors)
```

It is necessary to rename the column names of the dataset to ensure consistency. [Reference](https://www.datanovia.com/en/lessons/rename-data-frame-columns-in-r/).
```{r renaming the column names}
cocoa_flavors <- cocoa_flavors %>% 
  rename(
    company = Company...Maker.if.known.,
    bean_origin = Specific.Bean.Origin.or.Bar.Name,
    review_date = Review.Date,
    cocoa_percent = Cocoa.Percent,
    company_location = Company.Location,
    rating = Rating,
    bean_type = Bean.Type,
    broad_bean_origin = Broad.Bean.Origin
  )

# viewing the changes once more.
colnames(cocoa_flavors)
```
Next will be confirmig the data types of the columns.
```{r data types}
str(cocoa_flavors)
```
Converting the review_date type to date.
```{r review_date to date}
cocoa_flavors$review_date <- as.Date(as.character(cocoa_flavors$review_date), format = "%Y")
cocoa_flavors$review_date <- format(as.Date(cocoa_flavors$review_date), "%Y")
```


Viewing the structure once more.
```{r structure after changes}
str(cocoa_flavors)
```
## 4. Analysis.
The analysis is done objective wise.

*1. The coffee manufacturing companies with the highest ratings.*

To solve this, a rating of 3.5 and above is considered a high rating, while a rating of belo 3.5 is considered a low rating.
```{r companies with highest ratings}
cocoa_flavors %>% 
  filter(rating >= 3.5) %>% 
  group_by(company) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  arrange(desc(mean_rating), n=10)
```


*2. Year with maximum highest rating.*
```{r year with max rating}
cocoa_flavors %>% 
  filter(rating >= 3.5) %>% 
  group_by(review_date) %>% 
  summarize(mean_rating = mean(rating)) %>% 
  arrange(desc(mean_rating))
```


*3. Broad bean with highest rating*
```{r broad bean with highest mean rating}
cocoa_flavors %>% 
  filter(rating>=3.5) %>% 
  group_by(broad_bean_origin) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  arrange(desc(mean_rating))
```


*4. Countries of origin for companies with highest ratings.*
```{r countries with best ratings}
cocoa_flavors %>% 
  filter(rating>=3.5) %>% 
  group_by(company_location) %>% 
  summarise(mean_rating = mean(rating)) %>% 
  arrange(desc(mean_rating))
```



## 5. Share.
In this phase, I will create visualizations to best share the results of my analysis. The data is exported for visualizations in Tableau.

